An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-16 EconGrowth.Rnw
Now using nlmrt to do CES curve fits.

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
# tikz allows use of LaTeX formatting and font in graphs. Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
opts_chunk$set(
  # ********************************************************************************************************
  eval=FALSE,
  # ********************************************************************************************************
  warning=FALSE, #Tells whether to show warnings in the output.
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=FALSE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

@@

<<CES functions, eval=TRUE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish guess values for phi beta, zeta, lambda_L and lambda_E.
  sigmaGuess <- 0.663 # 0.663 for Sweden
  # gamma_EGuess <- 0.147 # 0.147 for Sweden
  gamma_EGuess <- 0.01
  betaGuess <- 0.7 # 0.7 for Sweden
  lambda_LGuess <- 0.011 # 0.011 for Sweden
  lambda_EGuess <- 0.018 # 0.018 for Sweden
  # Runs a non-linear least squares fit to the data with constraints
  modelCES <- nls(iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
                                                  * iCapStk^(1-beta) 
                                                  * iLabor^beta)^((sigma-1.0)/sigma) 
                          + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
                                                  * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0)), 
                  data=dataTable,
                  start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
                             lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
                  control=nls.control(maxiter=10000, tol=1e-06, minFactor=1/1024, 
                                      printEval=FALSE, warnOnly=TRUE),
                  algorithm = "port",
                  lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf),
                  upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf),
                  )
  return(modelCES)
}
print("US Q")
print(summary(cesModel("US", "Q")))

@@

The CES model is of the form

@@

\noindent where $e$ is any of $q$, $x$, or $u$; $\gamma_E$, $\lambda_L$, $\lambda_E$, $\beta$, and $\sigma$ are fitting parameters; and $\mathrm{e}$ is the base of the natural logarithm. Ths sum $\alpha + \beta$ is constrained to unity.

@@

The LINEX model is of the form

>>>>

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
# tikz allows use of LaTeX formatting and font in graphs. Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
opts_chunk$set(
  # ********************************************************************************************************
  eval=FALSE,
  # ********************************************************************************************************
  warning=FALSE, #Tells whether to show warnings in the output.
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=FALSE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

@@

<<CES functions, eval=TRUE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish default guesses for sigma, gamma_E, beta, lambda_L and lambda_E.
  sigmaGuess <- 0.663 # 0.663 for Sweden
  gamma_EGuess <- 0.01 # Minimum contribution by energy is the default.
  betaGuess <- 0.7 # 0.7 for Sweden
  lambda_LGuess <- 0.011 # 0.011 for Sweden
  lambda_EGuess <- 0.018 # 0.018 for Sweden
  lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf)
  upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf)
  if (countryAbbrev == "US"){
    gamma_EGuess <- 0.147 # 0.147 for Sweden
  }
  if (countryAbbrev == "UK"){
    sigmaGuess <- 0.5
    gamma_EGuess <- 0.147 # 0.147 for Sweden
    betaGuess <- 0.9
    lambda_LGuess <- 0.0
    lambda_EGuess <- 0.0
    lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-0.3, lambda_E=-0.3)
    upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= 0.3, lambda_E= 0.3)
  }
  if (countryAbbrev == "CN"){

  }
  # Runs a non-linear least squares fit to the data with constraints
#   modelCES <- nls(iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
#                                                   * iLabor^beta
#                                                   * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
#                           + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
#                                                   * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0)), 
#                   data=dataTable,
#                   start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
#                              lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
#                   control=nls.control(maxiter=10000, tol=1e-05, minFactor=1/1024, 
#                                       printEval=TRUE, warnOnly=TRUE),
#                   algorithm = "port",
#                   lower = lower,
#                   upper = upper
#                   )
  model <- iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
                                                  * iLabor^beta
                                                  * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
                          + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
                                                  * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))
  modelCES <- nlxb(model,
                   data=dataTable,
                   start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
                              lambda_L=lambda_LGuess, lambda_E=lambda_EGuess)
                   )
print(class(modelCES))
print(modelCES)
  modelCES <- wrapnls(modelCES)
print(class(modelCES))
  return(modelCES)
}
print("US Q")
print(summary(cesModel("US", "Q")))

@@

The CES model has the form

@@

\noindent where $e$ is any of $q$, $x$, or $u$; $\sigma$, $\gamma_E$, $\beta$, $\lambda_L$, and $\lambda_E$ are fitting parameters; and $\mathrm{e}$ is the base of the natural logarithm. Ths sum $\alpha + \beta$ is constrained to unity.

@@

The LINEX model has the form
